HyperLogLog: Exponentially Bad in Adversarial Settings

### 2021-05-05 docs/HLL_exp_bad/main.tex
removing last todo to make something that can be submitted
adding info on sim-based defns

In the original HLL proposal~\cite{hll}, relatively small and large cardinalities are handled separately to correct for systematic errors. When the raw cardinality estimate is below $\frac{5}{2}m$, the Hit Counting algorithm of Whang \emph{et al.}~\cite{hitcounting} is used to produce the final cardinality estimate. This estimator uses the number of empty buckets (whose register value is 0) instead of the number of leading zeros in the stream. A similar correction is applied for large cardinalities in order to take into account the likelihood of hash collisions. In addition, the cardinality estimate is usually rounded to a whole number before being output. The original HLL algorithm from~\cite{hll} is shown in Figure~\ref{fig:hll}. This algorithm is able to estimate cardinalities greater than $10^9$ with a typical standard error of 4\%, using only 1.5 kB of memory. Crucially, the proof of this~\cite[Theorem 1]{hll} makes the assumption that the inputs are drawn independently and uniformly at random from the input domain.\footnote{Technically, the results of~\cite{hll} are proven for a distribution obtained by making arbitrary replications and permutations after sampling independently and uniformly, to simulate input taken from an \emph{ideal multiset}. But they also hold without replication and permutation because HLL's internal state is invariant under repeated inputs.}

@@

\todo[inline]{Results from Redis or GBQ if we get them.}

@@

>>>>

In the original HLL proposal~\cite{hll}, relatively small and large cardinalities are handled separately to correct for systematic errors. When the raw cardinality estimate is below $\frac{5}{2}m$, the Hit Counting algorithm of Whang \emph{et al.}~\cite{hitcounting} is used to produce the final cardinality estimate. This estimator uses the number of empty buckets (whose register value is 0) instead of the number of leading zeros in the stream. A similar correction is applied for large cardinalities in order to take into account the likelihood of hash collisions. In addition, the cardinality estimate is usually rounded to a whole number before being output. The original HLL algorithm from~\cite{hll} is shown in Figure~\ref{fig:hll}. This algorithm is able to estimate cardinalities greater than $10^9$ with a typical standard error of 4\%, using only 1.5 kB of memory. Crucially, the proof of this~\cite[Theorem 1]{hll} makes the assumption that the inputs are drawn independently and uniformly at random from the input domain -- loosely, it is an average case analysis.\footnote{Technically, the results of~\cite{hll} are proven for a distribution obtained by making arbitrary replications and permutations after sampling independently and uniformly, to simulate input taken from an \emph{ideal multiset}. But they also hold without replication and permutation because HLL's internal state is invariant under repeated inputs.}

@@

%\todo[inline]{Results from Redis or GBQ if we get them.}

@@

Note also that, for our definition to meaningfully capture security against cardinality manipulation, it is important to restrict  ${\mathcal{S}}$ to doing random sampling when handling \texttt{insert($\cdot$)} queries. Otherwise, our definition could be vacuously satisfied (for example ${\mathcal{S}}$ could just sample the PRF key $K$ for itself and provide a perfect real-world view). This is somewhat different to common simulation-based security definitions, where ${\mathcal{S}}$'s input might be restricted (e.g.\ in a confidentiality definition, ${\mathcal{S}}$ might receive only $|m|$ in place of $m$) but its behaviour is not. Crucially, in our setting, security arises from inspecting the details of ${\mathcal{S}}$ and observing how it instantiates HLL in the average case.
